---
title: "CCM project"
author: "Klara Raiber"
date: "2024-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
# Three data sets: 1) Wave 13 of UK Household Longitudinal Study - Jan 2021 to May 2023; 2) UK Household Longitudinal Study wave overarching data needed for migration; 3) Annual Population Survey (APS) for information on people working in health sector also including migration (2022) by ONS

```{r eval=FALSE}

setwd("C:/Users/U679219/OneDrive - Radboud Universiteit/RU-Drive/Documents/Oxford 2024/prediction paper")

#libraries
library(haven)
library(dplyr)
library(readr)
library(ggplot2)


#read the Stata .dta file
raw_wave13 <- haven::read_stata("m_indresp.dta") #wave 13 of UK Household Longitudinal Study - Jan 2021 to May 2023; respondents data

table(raw_wave13$m_adla) # one example of ADL: Manage to walk stairs
table(raw_wave13$m_dvage) #age in years
table(raw_wave13$m_indinui_xw) # weights to be use

#migration data
raw_migration <- haven::read_stata("xwavedat.dta") # UK Household Longitudinal Study wave overarching data needed for migration data
table(raw_migration$generation) # migration generation - we want to know who is not born in the UK

#formal
raw_formal  <- haven::read_stata("apsp_jd22_eul_pwta22.dta") # Annual Population Survey (APS) for information on people working in health sector also including migration (2022) by ONS https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/methodologies/annualpopulationsurveyapsqmi

```

## health sector data
# Main variables: age, SIC codes (N: Health & social work; makes sense because caregivers can be seen as social workers?), CRYOX7_EUL_Main (country of birth)
```{r}
#Import information/main variables form the Annual Population Survey:
unique(raw_formal$AGE) # age in years 
unique(raw_formal$IN0792SM) # SIC codes
unique(raw_formal$CRYOX7_EUL_Main) # country of birth => not UK means migration
table(raw_formal$COUNTRY) # Country within UK; we select England = 1 and Wales = 2

#build life table with raw data
variables_for_life_table_formal <- raw_formal %>% 
  filter(COUNTRY == 1 | COUNTRY == 2) %>% # to select only England and Wales
  select(AGE, CRYOX7_EUL_Main, IN0792SM) %>%  # SIC main job 
    mutate(
    age = case_when(AGE == -9 ~ NA, 
                         AGE == -8 ~ NA,
                    AGE < 16 ~ NA,
                         TRUE ~ AGE), #rest same
    age = as.numeric(age)
  ) %>% 
  mutate(
    migrant =  case_when(CRYOX7_EUL_Main == 1  ~ 0,
                        CRYOX7_EUL_Main != 1  ~ 1),
    migrant = case_when(CRYOX7_EUL_Main == -8  ~ NA,
                        TRUE ~ migrant),
    migrant = as.numeric(migrant)) %>% 
  mutate(
    sic_health = case_when(IN0792SM ==  14 ~ 1,
                         IN0792SM != 14 ~ 0), 
    sic_health = case_when(IN0792SM == -9 ~ NA,
                          IN0792SM == -8 ~ NA,
                          TRUE ~ sic_health),
    sic_health = as.numeric(sic_health)) %>% 
    mutate(
      migrant_in_health = if_else((migrant == 1 & sic_health == 1), 1, 0)
    )


variables_for_life_table_formal <- na.omit(variables_for_life_table_formal)

#check
table(variables_for_life_table_formal$age)
table(variables_for_life_table_formal$migrant)
table(variables_for_life_table_formal$sic_health)
table(variables_for_life_table_formal$migrant_in_health)


#combine them in one life table - probably there is a nicer way to do this - but it works:
mytable <- table(variables_for_life_table_formal$age, variables_for_life_table_formal$migrant_in_health) # frequency table migration by age
table_age_migration <- data.frame(mytable) # add frequencies to a data frame

table_age_migration <- subset(table_age_migration, table_age_migration$Var2 == 1) # we are interested in migrants (1. so Var2)
table_age_migration <-  subset(table_age_migration, select = -Var2) # excluded unneeded information
#rename for clarity
names(table_age_migration)[names(table_age_migration) == 'Var1'] <- 'age'
names(table_age_migration)[names(table_age_migration) == 'Freq'] <- 'number of migrants'


#same for SIC classification:

mytable2 <- table(variables_for_life_table_formal$age, variables_for_life_table_formal$sic_health) # frequency table migration by SIC (health sector or not)
table_age_sector <- data.frame(prop.table(mytable2, 1)) # add frequencies to a data frame
table_age_sector <- subset(table_age_sector, table_age_sector$Var2 == 1) # we are interested in health workers (1. so Var2)
table_age_sector <-  subset(table_age_sector, select = -Var2)  # excluded unneeded information
#rename for clarity
names(table_age_sector)[names(table_age_sector) == 'Var1'] <- 'age'
names(table_age_sector)[names(table_age_sector) == 'Freq'] <- 'share_formal'


table_age_sector_absolut <- data.frame(mytable2) # add frequencies to a data frame
table_age_sector_absolut <- subset(table_age_sector_absolut, table_age_sector_absolut$Var2 == 1) # we are interested in health workers (1. so Var2)
table_age_sector_absolut <-  subset(table_age_sector_absolut, select = -Var2)  # excluded unneeded information
#rename for clarity
names(table_age_sector_absolut)[names(table_age_sector_absolut) == 'Var1'] <- 'age'
names(table_age_sector_absolut)[names(table_age_sector_absolut) == 'Freq'] <- 'number of people working in health'


#put them together in one life table:
life_table_formal <- table_age_sector %>% 
    inner_join(table_age_migration, by = "age")

life_table_formal <- life_table_formal %>% 
    inner_join(table_age_sector_absolut, by = "age")

life_table_formal$share_formal_mig <- life_table_formal$`number of migrants` / life_table_formal$`number of people working in health`

life_table_formal <-  subset(life_table_formal, select = -`number of migrants`) #get rid of unneeded information 
life_table_formal <-  subset(life_table_formal, select = -`number of people working in health`) # get rid of unneeded information 



```
# Information from understanding society
##Import information/main variables form the UKHLS (2023) for life tables: pidp (ID variable), m_aidhh (caregiving within household), m_aidxhh (caregiving outside household), to caculate age: m_istrtdaty & m_birthy, (I)ADLS: m_adla (manage stairs), m_adlb (get around house), m_adlc (get in/out bed), m_adld (cut toenails), m_adle (bath/shower), m_adlf (walk down the road), m_adlg (use toilet), m_adlh (eat including cutting food), m_adli (wash face and hands), m_adlj (dress/undress), m_adlk (take the right amount of medicine), m_adll (do shopping), m_adlm (do housework or laundry) & m_adln (doing paperwork or pay bills), individual 2022 cross-sectionals weights: m_indinui_xw, infromal help for care needs: m_hlpinfa96, formal help for care needs :m_hlpforma96
#From not wave specific information UKHLS: migration generation: generation

```{r}
unique(raw_wave13$m_adla)
       
variables_for_life_table_informal_need <- raw_wave13 %>% 
  filter(m_country == 1 | m_country == 2) %>% # to select only England and Wales
  select(pidp, m_aidhh, m_aidxhh, m_istrtdaty, m_birthy, 
         m_adla, m_adlb, m_adlc, m_adld, m_adle, m_adlf, 
         m_adlg, m_adlh, m_adli, m_adlj, m_adlk, m_adll, 
         m_adlm, m_adln, m_indinui_xw, m_hlpinfa96, m_hlpforma96) %>%  
    mutate(
    care_giving = if_else(m_aidhh == 1 | m_aidxhh == 1, 1, 0), #rest same
    care_giving = as.numeric(care_giving),
    care_giving_weighted = care_giving * m_indinui_xw
  ) %>% 
  mutate(
    age = m_istrtdaty - m_birthy,
    age = if_else(age > 101, NA, age)
  ) %>% 
  mutate(
    care_need = if_else(m_adla == 2 | m_adlb == 2 | m_adlc == 2 | m_adld == 2 
                        | m_adle == 2 | m_adlg == 2 | m_adlh == 2 | m_adli == 2 
                        | m_adlj == 2 | m_adlk == 2 | m_adla == 3 | m_adlb == 3 |
                          m_adlc == 3 | m_adld == 3 | m_adle == 3  | m_adlg == 3 
                        | m_adlh == 3 | m_adli == 3 | m_adlj == 3 |
                          m_adlk == 3 , 1, 0 ),
    care_need_weighted = care_need * m_indinui_xw) %>% 
  mutate(
    informal_received = as.numeric(if_else(m_hlpinfa96 == 0 & care_need == 1, 1, 0)), # not asked for f (walking down road), l (do the shopping), m (housework), n (do paperwork or pay bills) => reduce our care need to this definition (conservative estimate)
    informal_received_weighted = informal_received * m_indinui_xw
  ) %>% 
  mutate(
    formal_received = as.numeric(if_else(m_hlpforma96 == 0 & care_need == 1, 1, 0)),
    formal_received_weighted = formal_received * m_indinui_xw) %>%
  mutate(
    both_informal_formal_received = as.numeric(if_else(informal_received == 1 & formal_received == 1, 1, 0)),
    both_informal_formal_received_weighted = both_informal_formal_received * m_indinui_xw) %>% 
      select(pidp, care_giving, care_giving_weighted, age, care_need, care_need_weighted, 
             informal_received, informal_received_weighted, formal_received, 
             formal_received_weighted, both_informal_formal_received, 
             both_informal_formal_received_weighted, m_indinui_xw)


variables_for_life_table_informal_need_with_mig <- merge(variables_for_life_table_informal_need, raw_migration, by = 'pidp')  %>%  
  select(pidp, care_giving, care_giving_weighted, age, care_need, care_need_weighted, 
             informal_received, informal_received_weighted, formal_received, 
             formal_received_weighted, both_informal_formal_received, 
             both_informal_formal_received_weighted, m_indinui_xw, generation) %>% 
  mutate(
    migrant = as.numeric(if_else(generation == 1, 1, 0)),
    migrant_weighted = migrant * m_indinui_xw
  ) %>% 
  mutate(
    migrant_with_needs = as.numeric(if_else(migrant_weighted > 0 & care_need_weighted > 0, m_indinui_xw, 0)),
    migrant_give_care = as.numeric(if_else(migrant_weighted > 0 & care_giving_weighted > 0, m_indinui_xw, 0))
  )

variables_for_life_table_informal_need_with_mig <- na.omit(variables_for_life_table_informal_need_with_mig)

```

#Put it in a life table 
```{r}
unique(variables_for_life_table_informal_need_with_mig$age) # 16 to 101

#Empty data frame to fill
df <- data.frame(matrix(ncol = 17, nrow = 86))
x <- c('age', 'total_weight', 'need', 'share_need', 'care_giving', 'share_care_giving', 'migrant_need', 'total_migrant_weighted', 'share_need_mig', 'migrant_giving', 'share_informal_mig','sum_receiving_informal',  'share_with_care_needs_receiving_informal_care', 'sum_receiving_formal', 'share_with_care_needs_receiving_formal_care', 'sum_receiving_both', 'share_with_care_needs_receiving_both_care')
colnames(df) <- x

j <- 1

for (i in 16:101) {
   df <- df %>% 
mutate(age = replace(age, j, i),
     total_weight = replace(total_weight, j, (with(variables_for_life_table_informal_need_with_mig, sum(m_indinui_xw[age == i])))),
   need = (replace(need, j, (with(variables_for_life_table_informal_need_with_mig, 
                                  sum(care_need_weighted[age == i]))))),
  share_need = need / total_weight,
  care_giving = (replace(care_giving, j, (with(variables_for_life_table_informal_need_with_mig, 
                                  sum(care_giving_weighted[age == i]))))),
  share_informal = care_giving / total_weight,
  total_migrant_weighted = (replace(total_migrant_weighted, j,
                                    (with(variables_for_life_table_informal_need_with_mig, 
                                  sum(migrant_weighted[age == i]))))),
  migrant_need = (replace(migrant_need, j, (with(variables_for_life_table_informal_need_with_mig, 
                                  sum(migrant_with_needs[age == i]))))),
  share_need_mig = migrant_need / total_migrant_weighted,
  migrant_giving = (replace(migrant_giving, j, (with(variables_for_life_table_informal_need_with_mig, 
                                  sum(migrant_give_care[age == i]))))),
  share_informal_mig = migrant_giving / total_migrant_weighted,
  sum_receiving_informal = (replace(sum_receiving_informal, j,
                                                        (with(variables_for_life_table_informal_need_with_mig,
                                  sum(informal_received_weighted[age == i]))))),
  share_with_care_needs_receiving_informal_care = sum_receiving_informal / need,
  sum_receiving_formal = (replace(sum_receiving_formal, j,
                                                        (with(variables_for_life_table_informal_need_with_mig,
                                  sum(formal_received_weighted[age == i]))))),
  share_with_care_needs_receiving_formal_care = sum_receiving_formal / need,
  sum_receiving_both = (replace(sum_receiving_both, j,
                                                        (with(variables_for_life_table_informal_need_with_mig,
                                  sum(formal_received_weighted[age == i]))))),
  share_with_care_needs_receiving_both_care = sum_receiving_both / need,
   j = j + 1)
   }

life_table_part2 <- df %>%  select(age, share_need, share_informal, share_need_mig, share_informal_mig, share_with_care_needs_receiving_informal_care, share_with_care_needs_receiving_formal_care, share_with_care_needs_receiving_both_care)

#make age an integer 
life_table_formal$age <- as.integer(life_table_formal$age)

#add both life tables together
life_table_CCM_project <- life_table_part2 %>% 
    merge(life_table_formal, by = "age", all = TRUE)

#export to CSV
readr::write_csv(life_table_CCM_project, 
                 file = "life_table_CCM_project.csv")
#export is as R file
saveRDS(life_table_CCM_project, 
        file = "life_table_CCM_project.rds")

```
#Some descriptive graphs to look at the data

```{r}
#get the data
life_table_CCM_project <- readRDS("life_table_CCM_project.rds")

#graph on caregiving formal and informal
ggplot(life_table_CCM_project,aes(x = age, y = share_informal)) +
  geom_point() +
  xlim(16, 89) +
  theme_minimal() +
  labs(x = 'Age', y = 'Share providing informal care', title = 'Life table plotted')

#save 
ggsave("descrip_informal.png",bg="white")

ggplot(life_table_CCM_project,aes(x = age, y = share_formal)) +
  geom_point() +
  xlim(16, 65) +
  theme_minimal() +
  labs(x = 'Age', y = 'Share providing informal care', title = 'Life table plotted')

#save 
ggsave("descrip_formal.png",bg="white")

ggplot(life_table_CCM_project,aes(x = age, y = share_need)) +
  geom_point() +
  theme_minimal() +
  labs(x = 'Age', y = 'Share needing care', title = 'Life table plotted')

#save 
ggsave("descrip_need.png",bg="white")


```

